### 2016-06-16 capability-machine.tex
removed old malloc specification

The ticket dispenser test program uses $\malloc$. To be able to reason about programs that use $\malloc$, we assume that it obeys the following specification. If you provide it with one argument, namely the size of the allocation you wish and a return capability, then at some point $\pcreg$ will contain an executable capability based on this return capability and the return register will contain a $\rwx$ capability for the freshly allocated memory. 
\begin{specification}[Malloc (D)]
Take $W_{\malloc}$ the world with exactly one island, namely the unspecified island $\iota_{\malloc}$ that governs the internal state of $\malloc$.  If we invoke $\malloc$ in a memory $h = \hs_{\var{frame}} \uplus \hs_{\var{footprint}}$ that is the disjoint sum of a frame part and a footprint part and the latter satisfies this world up to steps $n$, i.e., $\heapSat[\hs_{\var{footprint}}]{n}{W_{\malloc}}$. Then $\malloc$ will succesfully terminate and return a result in a heap $h' = \hs_{\var{frame}} \uplus \hs_{\var{footprint}}'$ that is the disjoint sum of the same unmodified frame part and a potentially new footprint part, such that there is a future world $W'$ of $W_{\malloc}$ consisting of exactly two islands, where the first is an instance of $\iota_{\malloc}$ and the second is the region $\iota_{\var{base},\var{end}}$ for an appropriately-sized interval $\var{base},\var{end}$ and the new footprint part satisfies this new world $\heapSat[\hs_{\var{footprint}}']{?}{W'}$. \lau{at what step should this hold? Also it should probably be more precise what it means to ``invoke'' $\malloc$ and for $\malloc$ to ``return'' something.}
\end{specification}
\begin{specification}[Malloc (L)]
  For all worlds, $W$, heap segments, $\hs$, register files ,$\reg$, natural numbers, $n>j$ ($j$ is the number of steps the allocation runs for), and addresses $b$, $e$, and $return$ where $\heapSat[\hs]{n}{W}$ if
  \begin{itemize}
  \item $\memreg(r_0) = (\entry,b,e,\var{return})$,
  \item $\memreg(r_1) = m$ where $m>0$\footnote{We do not include 0 as malloc in C returns a null pointer when malloc(0) is called.}, and
  \item the remaining registers contain 0,
  \end{itemize}
then there exists addresses $b'$ and $e'$ where $b'-e' = m$ such that for all future worlds $W' \future W$ where $\heapSat[\hs']{n-j}{W'}$ for $\hs' = hs[b' \mapsto 0, \dots, e' \mapsto 0]$ we have $(\reg,\hs) \step[j] (\reg',\hs')$ \lau{need to extend $\hs$ to entire heap} for some register file $\reg'$ that has the following constraints:
\begin{itemize}
\item $\reg'(\pcreg) = (\rwx,b,e,\var{return})$,
\item $\reg'(r_0) = (\entry,b,e,\var{return})$,
\item $\reg'(r_1) = (\rwx,b',e',b')$, and
\item $\reg'(r) = 0$ for all other registers $r$.
\end{itemize}
\end{specification}

>>>>

The ticket dispenser test program uses $\malloc$. To be able to reason about programs that use $\malloc$, we assume that it obeys the following specification. If you provide it with one argument, namely the size of the allocation you wish and a return capability, then at some point $\pcreg$ will contain an executable capability based on this return capability and the return register will contain a $\rwx$ capability for the freshly allocated memory. 
\begin{specification}[Malloc (D)]
Take $W_{\malloc}$ the world with exactly one island, namely the unspecified island $\iota_{\malloc}$ that governs the internal state of $\malloc$.  If we invoke $\malloc$ in a memory $h = \hs_{\var{frame}} \uplus \hs_{\var{footprint}}$ that is the disjoint sum of a frame part and a footprint part and the latter satisfies this world up to steps $n$, i.e., $\heapSat[\hs_{\var{footprint}}]{n}{W_{\malloc}}$. Then $\malloc$ will succesfully terminate and return a result in a heap $h' = \hs_{\var{frame}} \uplus \hs_{\var{footprint}}'$ that is the disjoint sum of the same unmodified frame part and a potentially new footprint part, such that there is a future world $W'$ of $W_{\malloc}$ consisting of exactly two islands, where the first is an instance of $\iota_{\malloc}$ and the second is the region $\iota_{\var{base},\var{end}}$ \lau{doesn't this invariant allow the newly allocated memory to contain capabilities?} for an appropriately-sized interval $\var{base},\var{end}$ and the new footprint part satisfies this new world $\heapSat[\hs_{\var{footprint}}']{?}{W'}$. \lau{at what step should this hold? Just some step n'<n? Also it should probably be more precise what it means to ``invoke'' $\malloc$ and for $\malloc$ to ``return'' something.}
\end{specification}
